Using vector space models based on LDA and LSA for automatically grading exam questions

### 2018-06-18 tex/article.tex
Created a new section to discuss the other three variables (spelling correction, mapping algorithm and counting method) in more detail.

The LSA models on average did best with 100 SVs ($\textit{r}_s$ = .38) when being trained on the \textsc{student answers}, as compared to 20, 50 or 200 SVs. When training them on the \textsc{textbook chapter}, 10 SVs were used in the best average performance ($\textit{r}_s$ = .35), as compared to 5, 20, 50, 100, 200 or 400 SVs.

Thus, in training the final topic models, I set the number of topics or SVs to the above-reported numbers that had yielded the best outcomes on the two types of training data.

@@

\section{Discussion}
In this study I compared three types of vector space models for automatically predicting exam grades. The vector space models that used topics (LDA) or singular values (LSA) as their dimensions both outperformed the baseline vector space models that used vocabulary counts as their dimensions. The likely explanation for this is that LDA and LSA models compare documents at the semantic level, rather than looking for exact vocabulary matches \cite{kakkonen2008}. Notably, even in all of the baseline models the correlation between predicted and lecturer-assigned grades was statistically significant at the .05 level.

@@

\subsection{Are the best models good enough?}
I will now again zoom in on the best individual models, whose outcomes were visually represented in Figures \ref{lda}, \ref{lsa} and \ref{baseline}. Among other things, these figures showed that none of the best models ever predicted the grade of 10, even though this was the most common grade in the training set (see Figure 1). Thus, while the lecturer considered many of the students' answers to be excellent, these answers apparently were quite different from the reference answer in terms of the vector space models that I used for calculating semantic similarity. This shows that \citeasnoun{wolfe1998}'s reference answer approach may not be ideal for grading student answers by means of vector space models.

@@

\subsection{Suggestions for future research}
With regard to the three other variables, spelling correction and the mapping algorithm did not seem to have much of an effect in the baseline and LSA models. In the LDA models they did, but these models are the least reliable for evaluating these variables, because LDA models are probabilistic and their results were seen to fluctuate quite a bit when the same model was run twice. The baseline model worked best with raw counts; for LDA not much of a difference is seen, and for LSA it seems best to use TF-IDF counting.

The evaluation metric in this study was the correlation between the assigned and predicted grades. It would be interesting to consider other metrics too, for example the sum of squared errors (SSE) between the assigned and predicted grades. This would enable us to also evaluate a baseline model that would assign the most common grade (in this case 10) to all student answers. Such a baseline model currently could not be employed because a correlation coefficient cannot be calculated when there is no variance in the data.

>>>>

The LSA models on average did best with 100 SVs ($\textit{r}_s$ = .38) when being trained on the \textsc{student answers}, as compared to 20, 50 or 200 SVs. When training them on the \textsc{textbook chapter}, 10 SVs were used in the best average performance ($\textit{r}_s$ = .35), as compared to 5, 20, 50, 100, 200 or 400 SVs. Thus, in training the final topic models, I set the number of topics or SVs to the above-reported numbers that had yielded the best outcomes on the two types of training data.

@@

\newpage
\section{Discussion}
In this study I compared three types of vector space models for automatically predicting exam grades. The vector space models that used topics (LDA) or singular values (LSA) as their dimensions both outperformed the baseline vector space models that used vocabulary counts as their dimensions. The likely explanation for this is that LDA and LSA models compare documents at the semantic level, rather than looking for exact vocabulary matches \cite{kakkonen2008}. Notably, even in all of the baseline models the correlation between predicted and lecturer-assigned grades was statistically significant at the .05 level.

@@

\subsection{Spelling correction, mapping algorithm, and counting method}
Spelling correction and the mapping algorithm did not seem to have much of an effect in the baseline and LSA models. Possibly, the students did not make many spelling errors: Dutch students generally are quite proficient in English. Regarding the the mapping algorithm, this variable can only exert an effect on the outcomes when the models predict the grades 1 and 9 to begin with. This was often not the case (or, alternatively, very few 1's and 9's were predicted), which likely explains the lack of effect of this variable. Spelling correction and mapping algorithm did seem to have some effect in the LDA models (see Table \ref{table1}). However, among the three model types the LDA models are the least reliable for evaluating such variables, because they are probabilistic. In fact, their results were seen to fluctuate quite a bit when being trained multiple times with the same hyperparameter values.

Regarding the counting method, the baseline model worked best with raw counts; for LDA not much of a difference was seen, and for LSA it seems best to use TF-IDF counting, although the latter cannot be said with much certainty either due to the issue described in \ref{sec:training}. Therefore, the investigation of the influence of counting method should be considered only exploratory. To gain more insight in this variable, the computational issue should first be fixed. In addition, to obtain more reliable outcomes for the LDA models, it would be good if they were trained many times and their outcomes averaged.

\subsection{Are the best models good enough?}
I will now discuss the best individual models, whose outcomes were visually represented in Figures \ref{lda}, \ref{lsa} and \ref{baseline}. Among other things, these figures showed that none of the best models ever predicted the grade of 10, even though this was the most common grade in the training set (see Figure 1). Thus, while the lecturer considered many of the students' answers to be excellent, these answers apparently were quite different from the reference answer in terms of the vector space models that I used for calculating semantic similarity. This shows that \citeasnoun{wolfe1998}'s reference answer approach may not be ideal for grading student answers by means of vector space models.

@@

\subsection{Suggestions for future research}
The evaluation metric in this study was the correlation between the assigned and predicted grades. It would be interesting to consider other metrics too, for example the sum of squared errors (SSE) between the assigned and predicted grades. This would enable us to also evaluate a baseline model that would assign the most common grade (in this case 10) to all student answers. Such a baseline model currently could not be employed because a correlation coefficient cannot be calculated when there is no variance in the data.
